---
title: "Melanie_HW2"
author: "Bazgan"
date: "2026-02-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(ISLR)
numeric_vars <- Credit[, sapply(Credit, is.numeric)]

summary_stats <- data.frame(
  Variable = names(numeric_vars),
  Mean = sapply(numeric_vars, mean),
  Median = sapply(numeric_vars, median),
  Min = sapply(numeric_vars, min),
  Max = sapply(numeric_vars, max)
)

summary_stats

mean_balance <- mean(Credit$Balance)
median_balance <- median(Credit$Balance)

mean_balance
median_balance
```
2) If mean>median then it's a right/ positive skew
```{r}
par(mfrow = c(1,3))

hist(Credit$Income,
     main = "Histogram of Income",
     xlab = "Income (thousands)",
     breaks = 20)

hist(Credit$Balance,
     main = "Histogram of Balance",
     xlab = "Balance",
     breaks = 20)

hist(Credit$Rating,
     main = "Histogram of Rating",
     xlab = "Rating",
     breaks = 20)

par(mfrow = c(1,1)) 
```
4)
Income: positively skewed, a small number of high income customers create a right tail in the distribution.</br>
Balance: strongly positively skewed, many customers have low balances
and a few have very high balances creating noticeable outliers.</br>
Rating: Approximately symmetric to mildly right-skewed.Ratings have typical values with fewer extreme cases.</br>
```{r}
Credit$LogIncome <- log(Credit$Income)

par(mfrow = c(1,2))

hist(Credit$Income,
     main = "Original Income",
     xlab = "Income (thousands)",
     breaks = 20)

hist(Credit$LogIncome,
     main = "Log-Transformed Income",
     xlab = "log(Income)",
     breaks = 20)

par(mfrow = c(1,1))  
```
5) Compared to the original income histogram the right tail is compressed, the extreme high income values have less influence and the data is more evenly spread across the range.</br>
6) </br>
a-To explore the relationship between customer information and credit card balance, I would use scatterplots for numerical variables, boxplots for categorical variables, and correlation analysis to identify overall patterns.</br>
b-Based on the EDA, I would focus on credit limit, credit rating, income, and student status, since these variables show clear relationships with balance and help explain differences in how much customers owe.</br>
